---
title: "DA5030 Practice Problems 1"
author: "Jiaming Xu"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
---
```{r setup, include=FALSE}
library(tidyverse)
```

2.

```{r}
df <- read_csv('https://da5030.weebly.com/uploads/8/6/5/9/8659576/customertxndata.csv', col_names = F)
colnames(df) <- c("nvis", "ntrans", "os", "gender", "revenue")
df
```
3.

```{r}
summarise(df,
          total_revenue = sum(revenue, na.rm=T),
          mean_nvis = mean(nvis, na.rm=T),
          med_revenue = median(revenue, na.rm = T),
          sd_revenue = sd(revenue, na.rm = T),
          com_gender = names(sort(table(df$gender), decreasing=T))[1])
```
4.

```{r}
df %>% 
    select(gender, revenue) %>% 
    drop_na() %>% 
    ggplot() +
        geom_bar(mapping = aes( x = gender, y = revenue, fill = gender), stat = 'identity')
```

5.

Correlation between number of visits and customer costs is 0.739. This indicate a strong relationship between two variables.

```{r}
cor(df$nvis, df$revenue, method = "pearson", use = "pairwise.complete.obs")
```

6.

number of transactions and gender variables contain missing data. It's hard to infer missing data from other information since we have so few variables available. Therefore, I use the statistical value replacing the missing values. However, it's impossible to use statistical for gender since it's a categorical data, and I only apply this replacement to transactions.
I use "sapply" function to apply "sum(is.na( ))" function to columns in df to calculate number of missing values in each column.

```{r}
sapply(df, function(x) sum(is.na(x)))
```

7.

As the result in question 3, male is the mode of gender. It's easier to manually replace them since only two columns and they're different type of replacement.

```{r}
df$ntrans[is.na(df$ntrans)] <- round(mean(df$ntrans, na.rm = T), digits = 0)
df$gender[is.na(df$gender)] <- "male"
```

8.

generate a sequence same length as df, then turn it to 0 and 1s. Odd for 1 and even for 0.

```{r}
split_index <- seq(1, nrow(df)) %% 2
train_df <- df[split_index == 1, ]
valid_df <- df[split_index == 0, ]
```

9.

The difference between train and valid datasats is about $10, or about 2.2%. I guess this means different dataset will return different models evenif only switch purpose of two.

```{r}
print( sprintf( "Average revenue of train dataset: $%.2f, validation dataset: $%.2f. ",
                mean(train_df$revenue), mean(valid_df$revenue)))
```

10.

```{r}
set.seed(77654)
sample <- sample(1:3,size=nrow(df),replace=TRUE,prob=c(0.6,0.2,0.2))
train <- df[sample == 1, ]
valid <- df[sample == 2, ]
test <- df[sample == 3, ]
```



