---
title: "Practice 2"
author: "Jiaming Xu"
date: "1/29/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Q1:

Find the outliers of muders, 1.5 SD from the mean, displace information of outliers

```{r}
murder.mean <- mean(USArrests$Murder)
murder.sd <- sd(USArrests$Murder)
USArrests %>% 
  mutate( z = (Murder - murder.mean)/murder.sd) %>% 
  filter( abs(z) > 1.5)
```

## Q2:

plotting urban population and murder to visualize relationship. In the figure we can see there's no such a trend that murder increases as increasing urban population. Also the Pearson coefficient is very close to 0, which means there's nearly not correlation between urban population and murder.

```{r, echo = FALSE}
ggplot( data = USArrests) +
  geom_point( mapping = aes( x = UrbanPop, y = Murder))

print(cor( USArrests$UrbanPop, USArrests$Murder))
```

## Q3:

2-year weighted moving average:

```{r}
brazil <- read_csv("Brazil.csv")
pre1 <- brazil
pre1$pre_2yr <- 0
pre1$pre_2yr[1:2] = pre1$Subscribers[1:2]
pre1$pre_2yr[3: nrow(pre1)] = (pre1$Subscribers[2: (nrow(pre1) - 1)] * 5 +
                                   pre1$Subscribers[1: (nrow(pre1) - 2)] * 2) / 7
pre1$error <- pre1$Subscribers - pre1$pre_2yr
pre1$sqrt <- pre1$error^2
pre1
```

Exponential smoothing (alpha of 0.4):

```{r}
pre2 <- brazil
pre2$es <- 0
pre2$error <- 0
pre2$es[1] = pre2$Subscribers[1]
alpha <- 0.4

for (i in 2:nrow(pre2)) {
  pre2$es[i] <- pre2$es[i-1] + (alpha * pre2$error[i-1])
  pre2$error[i] <-  pre2$Subscribers[i] - pre2$es[i]
}
pre2$sqrt <- (pre2$error)^2
pre2
```

Linear regression trendline:

```{r}
pre3 <- brazil
pre3$lm <- 0
lm_coeff <- lm(pre3$Subscribers ~ pre3$Year)
pre3$lm <- lm_coeff$coefficients[2] * pre3$Year + lm_coeff$coefficients[1]
pre3$eror <-  pre3$Subscribers - pre3$lm
pre3$sqrt <- (pre3$eror)^2
pre3
```

## Q4:

```{r}
sqrt_errors <- tibble( sqer_2yr = sqrt(sum(pre1$sqrt)/ (length(pre1$sqrt[pre1$sqrt!=0]) - 1)), 
                  sqer_es = sqrt(sum(pre2$sqrt)/ (length(pre2$sqrt[pre2$sqrt!=0]) - 1)),
                  sqer_lm = sqrt(sum(pre3$sqrt)/ (length(pre3$sqrt[pre3$sqrt!=0]) - 1)))
sqrt_errors
```

Linear model has the smallest MSE.

## Q5:

Weighted average forecase:

```{r}
predict_summary <- bind_cols(pre1$Year, pre1$pre_2yr, pre2$es, pre3$lm)
colnames(predict_summary) <- c("Year", "yr", "es", "lm")
predict_summary$overall <- (4*predict_summary$yr + 2*predict_summary$es + predict_summary$lm)/7
predict_summary
```



















