---
title: "DA5020 Assignment II"
author: "Jiaming Xu"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(tidyverse)
```

Q1:

Dimension of mslaep is 83 x 11. Msleep is a tibble including 83 species' name, genus, vore, order, conservation, sleep and weight information, in total 11 variables (including name). First 5 variables are character class, last 5 are double class. Summary function returns amount of NAs in each group, sleep_rem, sleep_cycle, brainwt contain NAs.


```{r}
summary(msleep)
head(msleep)
```

Q2:

I delete NAs in "vore" group for better understanding how many observations are incomplete. From the bar chart we can see that nearly 50% animals were herbivore in this research, about 25% were carnivore and 25% were omnivore. about 6% were insectivore. 

```{r}
msleep %>% 
    select(vore) %>% 
    drop_na() %>% 
    ggplot() + 
        geom_bar( mapping = aes( x = vore, y = ..count../sum(..count..)*100, fill = vore)) +
        ylab("Percentage (%)")
```

Q3:

It looks like in this dataset, varnivores sleep longer than herbivores.

```{r}
msleep %>% 
    filter(vore == 'carni' | vore == 'herbi') %>%
    group_by( vore) %>%
    summarise( total_sleep_mean = mean(sleep_total))
```

Q4:

It looks like there's a strong linear correlation between log-log brain weight and body weight in all animals.

```{r}
ggplot( data = msleep, mapping = aes(x = bodywt, y = brainwt), na.rm = T) +
    geom_point( na.rm = T) +
    coord_trans(x = "log10", y = "log10")
```

Q5:

```{r}
cor(msleep$brainwt, msleep$bodywt, use = "pairwise.complete.obs", method="pearson")
```

0.934 means a very strong correlation between two variables, which confirms our result in Q4.

Q6:

I used z score to determine if an animal is outlier. There are 13 outliers in msleep dataset.

```{r}
msleep %>% 
    mutate( z = (sleep_total - mean(sleep_total))/sd(sleep_total)) %>% 
    filter( abs(z) > 1.5) %>% 
    select( name, z)
```

