---
title: "DA5020.A4.Jiaming.Xu"
subtitle: "Assignment 4"
author: "Jiaming Xu"
date: "2/7/2022"
output: html_document
---

##Q1:

According to data set dictionary, VendorID, RateCodeID, Payment_type, Trip_type are categorical variables, therefore conver them into factor.

1. Remove rows of pickup date and dropoff date both not in Feburary.

2. Drop ehail_fee column since only NA in it.

```{r}
library(tidyverse)
library(dplyr)
df <- read_csv("https://s3.amazonaws.com/nyc-tlc/trip+data/green_tripdata_2020-02.csv")
str(df)

df_clean <- df %>% 
    mutate_at( vars( VendorID, RatecodeID, payment_type, trip_type), as.factor) %>% 
    mutate( pick_up_month = format( lpep_pickup_datetime, "%Y-%b")) %>% 
    mutate( drop_off_month = format( lpep_dropoff_datetime, "%Y-%b")) %>% 
    filter(  !( drop_off_month != "2020-Feb" & pick_up_month != "2020-Feb")) %>% 
    dplyr::select( -ehail_fee)

df_clean
```

## Q2:

I made a function to find the most common value in a vector. Use match function to find the amount of unique values, then use which.max return it's maximum index, then extract it's value.

The most common hail type is street-hail, and most common payment method is credit card.

```{r}
findMode <- function(x) {
    u <- unique(x)
    return (u[which.max(tabulate(match(x, u)))])
}
findMode(df_clean$trip_type)
findMode(df_clean$payment_type)
```

## Q3:

In February, people like to hail a taxi on Thursday, Friday and Saturday. 

```{r}
df_clean %>% 
    mutate( date = format(lpep_pickup_datetime, "%d-%a")) %>% 
    group_by( date) %>% 
    count %>% 
    ggplot( )+
        geom_bar( mapping = aes( x = date, y = n),
                   fill='white', color = 'black', stat = 'identity') +
        labs( x = "Date in Feburary", y = 'Frequency', title = 'Hail taxi frequency in Feburary') +
    coord_flip()

```

## Q4:

Create a function that use str_match function in string package, use regex pattern " (\\d{2}):" and extract the second group inside the parentheses. Since str_match returns a matrix, remember use [, 2] to take out second column from result.

```{r}
library(stringr)
HourOfDay <- function(x){
    return (str_match(x, " (\\d{2}):")[, 2])
}

HourOfDay("2020-02-01 11:10:25")
```

## Q5 - Q6:

I combine Q5 and Q6 together so that I can use pipe to manipulate and visualize data continuously. From the figure below, it shows that the median of trip distance in February is larger at 5 and 6 am in the morning, taxi is high demanding for communicators!

```{r}
df_clean %>% 
    mutate( lpep_pickup_hour = HourOfDay(lpep_pickup_datetime)) %>% 
    group_by( lpep_pickup_hour) %>% 
    summarise( median_trip_distance = median(trip_distance)) %>% 
    ggplot() + 
        geom_bar( mapping = aes( x = lpep_pickup_hour, y = median_trip_distance)
                  , fill ='white', color = 'black', stat = 'identity') +
        labs( x = "Hour in a day", y = "Median trip distance")
```

## Q7:

I tried to use "count" function in pipe but failed to use filter function to find the maximum value, therefore I used summarise function to count in each group.

People hailed taxi on valentine's day the most on February!

```{r}
df_clean %>% 
    mutate(day = format( lpep_pickup_datetime, '%b-%d')) %>% 
    group_by(day) %>% 
    summarise(n = n()) %>% 
    filter( n == max(n))
```



